---
title: "Final Project: Music vs Mental Health Analysis"
author: "Calvin, Vy, Karen"
date: "December 2, 2023"
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Executive Summary

We aim to explore how music impact mental health by analyzing relevant data. This will help us determine how music is the most beneficial for mental health. We'll compare this to the amount of people suffering from certain mental conditions.

# Importing the dataset

```{r}
data <- read.csv('mxmh_survey_results.csv')
head(data)
```
```{r}
require('tidyr')
require('dplyr')
require('ggplot2')
```
# Making categories for Anxious, Depressed, and Insomniac

```{r}
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Anxious = ifelse(data$Anxiety > 5, 1, 0)
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Depressed = ifelse(data$Depression > 5, 1, 0)
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Insomniac = ifelse(data$Insomnia > 5, 1, 0)

head(data)
attach(data)
```

# Fitted model
```{r}
fitA <- lm(Hours.per.day ~ Age + Anxious + Depressed + Insomniac + Music.effects)
summary(fitA)
```
```{r}
fitB <- lm(Hours.per.day ~ Age + Anxious + Depressed + Insomniac + Music.effects)
summary(fitB)
```





# Non-linearity for the continuous variable
# Regression
# Correlation
# Normalities
# Looking at the significance in the dataset
# AIC (to see how well the model fits the data)
# Test violations (outliers and high leverage point)
# Analysis and Conclusion
# Code Appendix:

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```












