---
title: "Final Project: Music vs Mental Health Analysis"
author: "Calvin, Vy, Karen"
date: "December 2, 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width = 11, fig.height = 6)
```


# Executive Summary

We aim to explore how music impact mental health by analyzing relevant data. This will help us determine how music is the most beneficial for mental health. We'll compare this to the amount of people suffering from certain mental conditions.

# Importing the dataset

```{r, echo = FALSE}
data <- read.csv('mxmh_survey_results.csv')
head(data, 10)
```
```{r, echo = FALSE}
require('tidyr')
require('dplyr')
require('ggplot2')
```
# Making categories for Anxious, Depressed, and Insomniac

Since our data had people rate their anxiety, depression, insomnia, and OCD and a scale of 1 to 10 (only integers), we will make categories for `Anxiety`, `Depression`, `Insomnia`, and `OCD` so that the categories will be binary. For example, for `Anxiety`, we will have a category called `Anxious` that is 1 if `Anxiety` > 5 and 0 otherwise.
```{r, echo = FALSE}
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Anxious = ifelse(data$Anxiety > 5, 1, 0)
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Depressed = ifelse(data$Depression > 5, 1, 0)
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Insomniac = ifelse(data$Insomnia > 5, 1, 0)

head(data, 10)
attach(data)
```
\newpage
# Testing Non-linearity

We don't need to test for non-linearity for categorical variables so we will only test for non-linearity for the continuous variables, namely age.


```{r, echo = FALSE, warning = FALSE}
ggplot(pivot_longer(data = data, 2), aes(Hours.per.day, Age)) + 
  labs(title = 'Age vs. Hours per day') +
  theme_minimal() + geom_point()
```
There doesn't seem to be a linear relationship between age and hours per day. We can transform age by logging it so that the relationship looks less non-linear.
With log age:

```{r, echo = FALSE, warning = FALSE}
# log transform because there are a few high values and many low values
ggplot(pivot_longer(data = data, 2), aes(Hours.per.day, log(Age))) + labs(title = 'log(Age) vs. Hours per day') +
    theme_minimal() + geom_point()
```
```{r, echo = FALSE}
hist(Hours.per.day)
# Hours per day is not normal so we can log transform it or use gamma glm
```
Since the hours per day that people listen to music is not normally distributed, for each of the models, we will have a model where we log Hours.per.day and a model where we fit it to a log-link Gamma distribution.


Also, since half our models will have log Hours.per.day, we should look also look to see if log(Age) and log(Hours.per.day  + 1).

```{r, echo = FALSE, warning = FALSE}
# log transform because there are a few high values and many low values
ggplot(pivot_longer(data = data, 2), aes(log(Hours.per.day + 1), log(Age))) + labs(title = 'log(Age) vs. log(Hours per day + 1') +
    theme_minimal() + geom_point()
```
The relationship between log(Age) and log(Hours.per.day + 1) is not obviously non-linear.

\newpage
# Fitted model

Since the hours per day that people listen to music is not normally distributed, we created two fits, one where we log Hours.per.day and the other where we fit it to a log-link Gamma distribution.

```{r, echo = FALSE}
# fitA where we log Hours.per.day + 1 since Hours.per.day is not normally distributed.
fitAlog <- lm(log(Hours.per.day + 1) ~ log(Age) + Anxious + Depressed + Insomniac + Music.effects)
summary(fitAlog)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitAlog, which = c(1,2))
```

```{r, echo = FALSE}
# fitA where it's fitted to a gamma distribution
# For both fitAlog and fitAgamma, we add 1 to Hours.per.day because we can't log zero.
fitAgamma <- glm(Hours.per.day + 1~ log(Age) + Anxious + Depressed + Insomniac + Music.effects,
                 family = Gamma (link = 'log'))
summary(fitAgamma)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitAgamma, which = c(1,2))
```
Analysis:

  - For Log: From looking at the Residual Vs Fitted plot for fitA_log, it seems that the normality assumption is not violated because the shape of the fit is cloud-like without any noticeable pattern. This means that fitA_log does have constant variance. However, there are possible outliers such as point 696 or 19. For the Normal Q-Q plot, normality doesn't seem to be violated because most error points remains on the normality line. However, there are still evidences of outliers. 

  - For Gamma: Since the distribution is Gamma, it is possible to have cluster up negatively value in Residuals and Fitted. This means that fitA_gamma, doesn't seem to violating normality assumption. In another word, fitA_gamma have a constant Variance. However, there are possible outliers such as points 696 or 19. For Normal Q-Q, Gamma violated normality assumption because errors points are going off the normal line.

```{r, echo = FALSE}
# log Hours.per.day +1
fitBlog <- lm(log(Hours.per.day + 1) ~ log(Age) + Depressed + Insomniac)
summary(fitBlog)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitBlog, which = c(1,2))
```

```{r, echo = FALSE}
# gamma glm
fitBgamma <- glm(Hours.per.day + 1 ~ log(Age) + Depressed + Insomniac,
                 family = Gamma (link = 'log'))
summary(fitBgamma)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitBgamma, which = c(1,2))
```
Analysis:

  - For Log: From looking at the Residual Vs Fitted plot for fitA_log, it seem that normality assumption is not violated because the shape of the fit is cloud-like without any noticeable pattern. This means that fitA_log does have constant variance. However, there is possible outliers such as point 696 or 19. For the Normal Q-Q plot, normality doesn't seem to be violated because most error points remains on the normality line. However, there is still evidences of outliers. 

  - For Gamma: Since the distribution is Gamma, it is possible to have cluster up negatively value in Residuals and Fitted. This means that fitA_gamma, doesn't seem to violating normality assumption. In another word, fitA_gamma have a constant Variance. However, there are possible outliers such as points 696 or 19. For Normal Q-Q, Gamma violated normality assumption because errors points are going off the normal line.

# AIC
```{r, echo = FALSE}
AIC(fitAlog)
AIC(fitBlog)

AIC(fitAgamma)
AIC(fitBgamma)
```

# BIC
```{r, echo = FALSE}
BIC(fitAlog)
BIC(fitBlog)

BIC(fitAgamma)
BIC(fitBgamma)
```
fitAlog is a better fit than fitBlog according to AIC because it has a lower AIC.
fitBlog is a better fit than fitAlog according to BIC because it has a lower BIC.

fitAgamma is a better fit than fitBgamma according to AIC because it has a lower AIC.
fitBgamma is a better fit than fitAgamma according to BIC because it has a lower BIC.



# Analysis and Conclusion




\newpage
# Code Appendix:

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```












