---
title: "Final Project: Music vs Mental Health Analysis"
author: "Hieu (Calvin) Hoang, Vy Vo, Karen Nguyen"
date: "December 2, 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align='center')
```

# Distribution Report: 

Hieu (Calvin) Hoang: Organization, decoding, interpretation/analysis and editor
Karen Nguyen: Coding, interpretation/analysis, and writer
Vy Vo: Coding, interpretation/analysis, and writer

# Introduction

In this project, we will investigate the relationship between how many hours per day someone listens to music (our response variable) and their age and mental health. To do this, we used [data collected from a music and mental health survey](https://www.kaggle.com/code/laurenlavelle/mental-health-and-music-analysis/input).

# Importing the dataset

```{r, echo = FALSE, message = FALSE}
require('tidyr')
require('dplyr')
require('ggplot2')
```

```{r, echo = FALSE, results = FALSE}
data <- read.csv('mxmh_survey_results.csv')
data <- data %>% drop_na(Hours.per.day, Age, Anxiety, Depression, Insomnia, Music.effects)
head(data, 10)
```

# Making categories for Anxious, Depressed, and Insomniac

Since our data had people rate their anxiety, depression, insomnia, and OCD and a scale of 1 to 10 (only integers), we will make categories for `Anxiety`, `Depression`, `Insomnia`, and `OCD` so that the categories will be binary. For example, for `Anxiety`, we will have a category called `Anxious` that is 1 if `Anxiety` > 5 and 0 otherwise.
```{r, echo = FALSE, results = FALSE}
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Anxious = ifelse(data$Anxiety > 5, 1, 0)
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Depressed = ifelse(data$Depression > 5, 1, 0)
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Insomniac = ifelse(data$Insomnia > 5, 1, 0)

head(data, 10)
attach(data)
```
# Testing Non-linearity

We don't need to test for non-linearity for categorical variables so we will only test for non-linearity for the continuous variables, namely age.

## Plots
```{r, echo = FALSE, warning = FALSE}
ggplot(pivot_longer(data = data, 2), aes(Hours.per.day, Age)) + 
  labs(title = 'Age vs. Hours per day') +
  theme_minimal() + geom_point()
```
There doesn't seem to be a linear relationship between age and hours per day. We can transform age by logging it so that the relationship looks less non-linear.
With log age:

```{r, echo = FALSE, fig.height = 4}
# log transform because there are a few high values and many low values
ggplot(pivot_longer(data = data, 2), aes(Hours.per.day, log(Age))) + labs(title = 'log(Age) vs. Hours per day') +
    theme_minimal() + geom_point()
```
```{r, echo = FALSE, fig.height = 4}
hist(Hours.per.day)
# Hours per day is not normal so we can log transform it or use gamma glm
```
Since the hours per day that people listen to music is not normally distributed, for each of the models, we will have a model where we log Hours.per.day and a model where we fit it to a log-link Gamma distribution.


Also, since half our models will have log Hours.per.day, we should look also look to see if log(Age) and log(Hours.per.day  + 1).

```{r, echo = FALSE, fig.height = 4}
# log transform because there are a few high values and many low values
ggplot(pivot_longer(data = data, 2), aes(log(Hours.per.day + 1), log(Age))) + labs(title = 'log(Age) vs. log(Hours per day + 1') +
    theme_minimal() + geom_point()
```
The relationship between log(Age) and log(Hours.per.day + 1) is not obviously non-linear, so we can use log(Age) in the linear model with log(Hours.per.day + 1).


# Fitted models

For fitA, which will have a log model, `fitAlog`, with covariates `log(Age)`,`Anxious`, `Depressed`, `Insomniac`, and `Music.effects`.  
`Anxious`, `Depressed`, and `Insomniac` are the binary categories we made earlier.  
`Music.effects` is a categorical variable where people reported what effect they felt music had on their mental health. The categories for `Music.effects` are `Improve`, `No effect`, and `Worsen`.

```{r, echo = FALSE, fig.width = 11, fig.height = 6, fig.cap = "Residual vs. Fitted Values Plot"}
## Fits

# fitA where we log Hours.per.day + 1 since Hours.per.day is not normally distributed.
fitAlog <- lm(log(Hours.per.day + 1) ~ log(Age) + Anxious + Depressed + Insomniac + Music.effects)
summary(fitAlog)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitAlog, which = 1, caption = '', main = 'Residual vs. Fitted Values Plot', col = 'lightskyblue')

plot(fitAlog, which = 2, caption = '', main = 'Q-Q Plot', col = 'lightskyblue')

```

```{r, echo = FALSE, fig.width = 11, fig.height = 6}
# fitA where it's fitted to a gamma distribution
# For both fitAlog and fitAgamma, we add 1 to Hours.per.day because we can't log zero.
fitAgamma <- glm(Hours.per.day + 0.001 ~ log(Age) + Anxious + Depressed + Insomniac + Music.effects,
                 family = Gamma (link = 'log'))
summary(fitAgamma)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitAgamma, which = 1, caption = '', main = 'Residual vs. Fitted Values Plot', col = 'lightskyblue')

plot(fitAgamma, which = 2, caption = '', main = 'Q-Q Plot', col = 'lightskyblue')


```
Analysis:

  - For Log: From looking at the Residual Vs Fitted plot for fitAlog, it seems that the normality assumption is not violated because the shape of the fit is cloud-like without any noticeable pattern. This means that fitA_log does have constant variance. However, there are possible outliers such as point 695 or 18. For the Normal Q-Q plot, normality doesn't seem to be violated because most error points remains on the normality line. Nevertheless, there are still evidences of outliers. 

  - For Gamma: Since the distribution is Gamma, it is possible to observe clustering of negatively points in Residuals and Fitted. This means that fitAgamma, doesn't seem to violate normality assumption because there is no pattern. In another word, fitAgamma does appear to have a constant Variance. However, there are possible outliers such as points 696 or 19. For the Normal Q-Q plot, Gamma violated normality assumption, as errors points are going off the normal line.

Overall Analysis:
Evidence from the Residual vs. Fitted plots and Q-Q plots for FitA and FitB indicates that the log model performs better than the gamma model as it doesn't violate normality. Therefore, we've chosen to exclude the gamma model and proceed with analyzing the two log fits.
  
For the fitB's, `fitBlog` and `fitBgamma`, we used only the covariates that were individually significant in `fitA` and `fitAgamma`. As a result, the covariates for `fitBlog` and `fitBgamma` are `log(Age)`, `Depressed`, and `Insomniac`.

```{r, echo = FALSE, fig.width = 11, fig.height = 6}
# log Hours.per.day + 1
fitBlog <- lm(log(Hours.per.day + 1) ~ log(Age) + Depressed + Insomniac)
summary(fitBlog)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitBlog, which = 1, caption = '', main = 'Residual vs. Fitted Values Plot', col = 'lightskyblue')

plot(fitBlog, which = 2, caption = '', main = 'Q-Q Plot', col = 'lightskyblue')


```

```{r, echo = FALSE, fig.width = 11, fig.height = 6}
# gamma glm
fitBgamma <- glm(Hours.per.day + 0.001 ~ log(Age) + Depressed + Insomniac,
                 family = Gamma (link = 'log'))
summary(fitBgamma)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitBgamma, which = 1, caption = '', main = 'Residual vs. Fitted Values Plot', col = 'lightskyblue')

plot(fitBgamma, which = 2, caption = '', main = 'Q-Q Plot', col = 'lightskyblue')


```
Analysis:

  - For Log: From looking at the Residual Vs Fitted plot for fitBlog, it seems that the normality assumption is not violated because the shape of the fit is cloud-like without any noticeable pattern. This means that fitB_log does have constant variance. However, there are possible outliers such as point 695 or 18. For the Normal Q-Q plot, normality doesn't seem to be violated because most error points remains on the normality line. Nevertheless, there are still evidences of outliers.

  - For Gamma: Since the distribution is Gamma, it is possible to observe clustering of negatively valued residuals in Residuals and Fitted. From the plot, it is evident that Gamma violates the normality assumptions. 

Overall Analysis:
Evidence from the Residual vs. Fitted plots and Q-Q plots for FitA and FitB indicates that the log model performs better than the gamma model as it doesn't violate normality. Therefore, we've chosen to exclude the gamma model and proceed with analyzing the two log fits.

# AIC
```{r, echo = FALSE, results = FALSE}
##AIC

AIC(fitAlog)
AIC(fitBlog)

```
AIC(fitAlog) = `r AIC(fitAlog)`   
AIC(fitBlog) = `r AIC(fitBlog)`  


# BIC
```{r, echo = FALSE, results = FALSE}
## BIC

BIC(fitAlog)
BIC(fitBlog)

```
BIC(fitAlog) = `r BIC(fitAlog)`   
BIC(fitBlog) = `r BIC(fitBlog)`  


# AIC and BIC analysis
fitAlog is a better fit than fitBlog according to AIC because it has a lower AIC.
fitBlog is a better fit than fitAlog according to BIC because it has a lower BIC.


Since which model is better according to AIC and BIC is different, we can use either fitA or fitB. One isn't clearly better than the other. We can confirm this with anova tests.

# ANOVA F-Tests
```{r}
## ANOVA tests

anova(fitAlog, fitBlog)

```
The null hypothesis $H_0$: $\beta_2 = \beta_5 = 0$ vs. $H_1$: $\beta_2 \ne 0$ or $\beta_2 \ne 0$. We fail to reject the null because the $Pr(>F) = 0.06215 > \alpha = 0.05$. As a result, fitAlog and fitBlog are the same so we will use the smaller model, fitBlog.

# Interpretation of the final model
Since the gamma distributions violate normality, we would prefer not to use them for our final model. Based on the AIC, BIC, and ANOVA models, which show us that `fitAlog` and `fitBlog` are about the same, we will use the smaller model for our final model.
As a result, our final model is `fitBlog`, which is:
$log(Hours.per.day + 1) =  1.785 -0.168 log(Age) + 0.133 Depressed + 0.115 Insomniac$

According to `fitBlog`: 
- If log(Age) increase by 1 unit then log(Hours.per.day + 1) will decrease by 0.168 units.
- If a person ranks their Depression above 5 on a scale from 1 to 10, then log(Hours.per.day + 1) will increase by 0.133 units.
- If a person ranks their Insomnia above 5 on a scale from 1 to 10, then log(Hours.per.day + 1) will increase by 0.115 units.

One thing to note for our models is that our $R^2$'s were low, including fitBlog, which adjusted $R^2 = 0.0397$. We realize this means that our models, including fitBlog, explain very little variation in the response (transformed hours per day).

Overall, there 

# Discussion

In our project, we faced a few minor setbacks. For example, our dataset contained a notable amount of missing values (NA), requiring us to remove rows to ensure accurate calculations for AIC, BIC, and ANOVA. By removing only a small portion of the data, we understand that it could have potentially affected our calculation.

Another difficulty we encountered was the large amount of categorical variables in our dataset. While categorical variables are useful and we did use many in our model, more quantitative variables would be helpful. 

Finally, we attempted to find predicted MSE by using LOOCV and kfoldCV to further support the best-fit model argument. However, despite consulting the TA, we continuously encountered errors indicating differences in variable sizes. Ultimately, we decided that since we already have results from ANOVA, AIC and BIC, it is not worth continuing to debug. We firmly believe that the MSE results would align with our other tests, supporting that fitBlog is our best model.

# Conclusion

Overall, we decide that fitBlog is the best model out of the four we made because it does best to capture the relationship between Hour Per Day with Age and Mental Health. 

This project have inspire a possible future study ideas where we utilize the many music genre variables that our dataset and possibly investigate how different music genre affect mental health. 


\newpage
\begin{center} Appendix: R Script \end{center}

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```












