---
title: "Final Project: Music vs Mental Health Analysis"
author: "Calvin, Vy, Karen"
date: "December 2, 2023"
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Executive Summary

We aim to explore how different music genres impact mental health by analyzing relevant data. This will help us determine which genre is the most beneficial for mental health. We'll compare this to the amount of people suffering from certain mental conditions from each genre.

# Importing the dataset

```{r}
data <- read.csv('mxmh_survey_results.csv')
head(data)
```

# Libraries

```{r}
require('tidyr')
require('dplyr')
require('ggplot2')
```

# Analysis

```{r}
# Top music favorite genres
genre_counts <- table(mxmh_survey_results$Fav.genre)
sorted_genres <- sort(genre_counts, decreasing = TRUE)
sorted_genres
```

- As shown from the data above, the top 3 music that people listen to are Rock, Pop, Metal.

- Next, we'll be analyzing how these genres affect people who listen to them.

```{r}
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Anxiety = ifelse(data$Anxiety > 5, 1, 0)
head(data)
```

```{r}
sum(data$Music.effects == 'Improve') 
```

```{r}
# can investigate improvement in people who are not anxious + together if have room in paper
# i.e. with ratio 2 + not sure how to do together lol
#split data frame based on particular column value
data_anxious <- data[data$Anxious == 1, ]
data_not_anxious <- data[data$Anxious == 0, ]
```

```{r}
head(data_anxious)
```



```{r}
rock_data <- data[data[["Fav genre"]] == "Rock", c("Fav genre", "Music effects", "Anxiety", "Depression")]
rock_indices <- which(data[["Fav genre"]] == "Rock")
rock_data <- data[data[["Fav genre"]] == "Rock", c("Fav genre", "Music effects", "Anxiety", "Depression")]
rock_data$Index <- rock_indices
#print(rock_data)
tail(rock_data)
```

```{r}
# CORRECT ratio1
ratio1 <- sum(data_anxious$Music.effects == 'Improve') / sum(data$Anxious == 1)
ratio1
```

```{r}
# For the first part, only include people with anxiety > 5 so the ratio includes people who's mental health improved divided by number of people who are anxious
```












