---
title: "Project Try 2"
author: "Karen Nguyen"
date: "2023-12-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, results = FALSE}
# Importing the dataset

data <- read.csv('C:/Users/karen/Documents/STA141A/mxmh_survey_results.csv')
attach(data)
head(data)
```

```{r, echo = FALSE, message = FALSE}
# Libraries

require('tidyr')
require('dplyr')
require('ggplot2')
```

# Making categories for Anxiety, Depression, Insomnia, and OCD.

Since our data had people rate their anxiety, depression, insomnia, and OCD and a scale of 1 to 10 (only integers), we will make categories for `Anxiety`, `Depression`, `Insomnia`, and `OCD` so that the categories will be binary. For example, for `Anxiety`, we will have a category called `Anxious` that is 1 if `Anxiety` > 5 and 0 otherwise.
```{r, echo = FALSE, results= FALSE, message=FALSE}
# mark people as anxious (1) if anxiety > 5, not anxious (0) otherwise
data$Anxious = ifelse(Anxiety > 5, 1, 0)
# mark people as depressed (1) if depression > 5, not depressed (0) otherwise
data$Depressed = ifelse(Depression > 5, 1, 0)
# mark people as insomniac (1) if insomnia > 5, not insomniac (0) otherwise
data$Insomniac = ifelse(Insomnia > 5, 1, 0)
# mark people as having ocd (1) if ocd > 5, not having ocd (0) otherwise
data$ocd = ifelse(OCD > 5, 1, 0)

head(data)
attach(data)
```
# Testing Non-linearity

We don't need to test for non-linearity for categorical variables so we will only test for non-linearity for the continuous variables, namely age.
```{r, echo = FALSE, warning = FALSE}
ggplot(pivot_longer(data = data, 2), aes(Hours.per.day, Age)) + labs(title = 'Age vs. Hours per day') +
    theme_minimal() + geom_point()
```
There doesn't seem to be a linear relationship between age and hours per day. We can transform age by logging it so that the relationship looks less non-linear.
With log age:
```{r, echo = FALSE, warning = FALSE}
# log transform because there are a few high values and many low values
ggplot(pivot_longer(data = data, 2), aes(Hours.per.day, log(Age))) + labs(title = 'log(Age) vs. Hours per day') +
    theme_minimal() + geom_point()
```
```{r, echo = FALSE}
hist(Hours.per.day)
# Hours per day is not normal so we can log transform it or use gamma glm
```
Since the hours per day that people listen to music is not normally distributed, for each of the models, we will have a model where we log Hours.per.day and a model where we fit it to a log-link Gamma distribution.

Also, since half our models will have log Hours.per.day, we should look also look to see if log(Age) and log(Hours.per.day  + 1).
```{r, echo = FALSE, warning = FALSE}
# log transform because there are a few high values and many low values
ggplot(pivot_longer(data = data, 2), aes(log(Hours.per.day + 1), log(Age))) + labs(title = 'log(Age) vs. log(Hours per day + 1') +
    theme_minimal() + geom_point()
```
The relationship between log(Age) and log(Hours.per.day + 1) is not obviously non-linear.

# Fitted models

```{r, echo = FALSE}
# DELETE LATER
# we can add more variables if we want to for A
# have as many fits as long as the models make sense 
# we can see if these fits are similar in analysis
```


```{r, echo = FALSE}
# fitA where we log Hours.per.day + 1 since Hours.per.day is not normally distributed.
fitAlog <- lm(log(Hours.per.day + 1) ~ log(Age) + Anxious + Depressed + Insomniac + Music.effects)
summary(fitAlog)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitAlog, which = c(1,2))
```

```{r, echo = FALSE}
# fitA where it's fitted to a gamma distribution
# For both fitAlog and fitAgamma, we add 1 to Hours.per.day because we can't log zero.
fitAgamma <- glm(Hours.per.day + 1~ log(Age) + Anxious + Depressed + Insomniac + Music.effects,
                 family = Gamma (link = 'log'))
summary(fitAgamma)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitAgamma, which = c(1,2))
```
fitB <- lm(Hours.per.day ~ log(Age) + Depressed + Insomniac)
summary(fitB)

```{r, echo = FALSE}
# log Hours.per.day +1
fitBlog <- lm(log(Hours.per.day + 1) ~ log(Age) + Depressed + Insomniac)
summary(fitBlog)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitBlog, which = c(1,2))
```

```{r, echo = FALSE}
# gamma glm
fitBgamma <- glm(Hours.per.day + 1 ~ log(Age) + Depressed + Insomniac,
                 family = Gamma (link = 'log'))
summary(fitBgamma)

# Residuals vs. Fitted Values Plot and QQ Plot
par(mfrow = c(1,2))
plot(fitBgamma, which = c(1,2))
```
# AIC
```{r, echo = FALSE}
AIC(fitAlog)
AIC(fitBlog)

AIC(fitAgamma)
AIC(fitBgamma)
```

# BIC
```{r, echo = FALSE}
BIC(fitAlog)
BIC(fitBlog)

BIC(fitAgamma)
BIC(fitBgamma)
```
fitAlog is a better fit than fitBlog according to AIC because it has a lower AIC.
fitBlog is a better fit than fitAlog according to BIC because it has a lower BIC.

fitAgamma is a better fit than fitBgamma according to AIC because it has a lower AIC.
fitBgamma is a better fit than fitAgamma according to BIC because it has a lower BIC.